---
title: "WUP_04_Alt"
author: "Kiara Butler!"
date: "09/14/2023"
output:
  html_document:
    df_print: paged
---

```{r include=FALSE,echo=FALSE}
require(knitr)
require(tidyverse)
require(tigerstats)
mydata <- read.csv(file="GSScsv.csv")

```

# Introduction

<!--In this section you explain what you are trying to show.  Where did the data come from?  What is the research or other question you are trying to answer?!-->

The data is in the file GSScsv.csv and came from the 2018 General Social Survey.  The question to answer is to resolve this hypothesis.  
$H_0$ There is no difference between the average score on the WORDSUM test and the RACE of the person taking the test   
$H_a$ There is a significant difference in average score on the WORDSUM test between at least two different RACES

# Methods

<!--Mention that you intend to use chatGPT to the greatest extent possible. Give credit to chatGPT when you use it to create code in your report. Decide on your methods:  use "variable analysis" or other appropriate descriptors.  Make sure to choose at least one graphical method and at least one numerical method.!-->

```{r eval=FALSE}
mydata <- mydata %>%
mutate(across(where(is.character), as.factor))
```

We have changed all the character variable the factor variables. This type of problem needs a numerical variable, WORDSUM and a factor variable called RACE.

# Results

<!--Divide this section into two sub-sections:  One for your descriptive  results and one for your inferential results.!-->

## Descriptive Results

### Graphical Descriptive Results

<!--Graphical results here.  Make sure to show your code.  Provide appropriate labels for axes, giving units if possible, and provide a good title for the graph, too.  Use the graphical results to describe the patterns if any that exist in the data as focused toward the research question!-->


```{r}
basicCN <- ggplot(mydata,aes(x=RACE,y=WORDSUM))

basicCN + geom_boxplot()
```

By looking at the OTHER RACE we see it's lower that the WHITE and BLACK RACE, and the WHITE RACE is higher than both, however we would need to do a statistical test to further test any difference.

### Numerical Descriptive Results

<!--Numerical results go here. Use the numerical results to describe the patterns if any that exist in the data as focused toward the research question!-->

```{r}
favstats(WORDSUM~RACE, data=mydata)
```

By looking at the mean we can see that race OTHER has a mean of 7.444444 while BLACK and WHITE have a similar mean with 8.730769 and 8.977612. Will need to run a test to see if the difference is significant.

## Inferential Results

<!--State hypothesis clearly.  Make sure your discussion of the inferential test covers all the aspects that the test output produces, such as test statistic, p-value etc.  Make a decision about the null hypothesis, explain the assumptions on which the selected test/procedure was based, and why the chosen procedure satisfys the assumptions and is appropriate to answer the research question!-->


```{r}
model1 <- aov(WORDSUM~RACE, data = mydata)
summary.aov(model1)
```
we got a signification result where  run a follow up multiple comparison test due to low p- value (.0182< .05)



In the multiple comparison test we can test we can see that in the 95% confidence, we can see  the dotted line that is indicate with a zero confidence level intersects WHITE-BLACK and OTHER-BLACK unlike WHITE-OTHER which is the reason why we result in a low p-value 
# Discussion and Conclusion

<!--Discussion and conclusion here.  If you found a relationship be sure to consider whether the relationship occurs because one of the variavbles causes the other, or whether they perhasps are related for some other reason.  Watch the chapter 6 videos from the GeorgeTown videos collection.!-->


In this write up we pretty much had to create a box plot for our factor value (RACE) and numerical value (WORDSUM) right from the beginning we noticed that there was an uncertainty of whether oer not there is any differences and needed to run more test. we then noticed somewhat of differences but not enough to come to a conclusion having us run anova test to see if we had a low p-value so we can then needed to run a multiple comparson test where we could finally see the difference and thus can come to conclusion 